HPC On-boarding

### 2016-01-19 hpc_101/paper.tex
Filled in section about slurm partitions.  Also created a partition layout image.

@@

\author{Sean Cleveland Ph.D, Ron Merrill Ph.D,\\David Schanzenbach M.S.} % Your name
\institute[University of {\hawaii} -- ITS--CI] % Your institution as it will appear on the bottom of every slide, may be shorthand to save space
{
Information Technology Services \\
Department of Cyberinfrastructure \\
University of {\hawaii} \\ % Your institution for the title page
\medskip
\textbf{\url{https://www.hawaii.edu/its/ci/}}\\
\textbf{\textit{uh-hpc-help@lists.hawaii.edu}} % Your email address
}
\date{\today} % Date, can be changed to a custom date

@@

\begin{frame}
	\frametitle{Overview -- Cluster Interaction}
	\begin{itemize}
		\item{Connecting to a cluster @ UH}
			\begin{itemize}
			\item Login to the cluster
			\item Verify user permissions
			\end{itemize}
    {\semitransp[25]{\item User directories
		\item Transferring files
		\begin{itemize}{\semitransp[25]{
			\item Globus
		}}\end{itemize}
		\item Software
		\begin{itemize}{\semitransp[25]{
			\item Modules
			\item Acquiring software
			\item Compiliers
			}}
		\end{itemize}
		\item Managing user jobs
		\begin{itemize}{\semitransp[25]{
			\item Job scheduler
			\item Using SLURM
			\item Partition layout			
			\item Submitting jobs (Examples)
			}}
		\end{itemize}			
		}}
	\end{itemize}
\end{frame}

@@

\begin{frame}
	\frametitle{Overview -- Cluster Interaction}
	\begin{itemize}
		\item {\semitransp[25]{Connecting to a cluster @ UH}}
			\begin{itemize}{\semitransp[25]{
			\item Login to the cluster
			\item Verify user permissions
			}}
			\end{itemize}
    \item User directories
		{\semitransp[25]{\item Transferring files
		\begin{itemize}{\semitransp[25]{
			\item Globus
		}}\end{itemize}
		\item Software
		\begin{itemize}{\semitransp[25]{
			\item Modules
			\item Acquiring software
			\item Compilers
			}}
		\end{itemize}
		\item Managing user jobs
		\begin{itemize}{\semitransp[25]{
			\item Job scheduler
			\item Using SLURM
			\item Partition layout			
			\item Submitting jobs (Examples)
			}}
		\end{itemize}			
		}}
	\end{itemize}
\end{frame}

@@

\begin{frame}
	\frametitle{Overview -- Cluster Interaction}
	\begin{itemize}
		\item {\semitransp[25]{Connecting to a cluster @ UH}}
			\begin{itemize}{\semitransp[25]{
			\item Login to the cluster
			\item Verify user permissions
			}}
			\end{itemize}
    {\semitransp[25]{\item User directories}}
		\item Transferring files
		\begin{itemize}
			\item Globus
		\end{itemize}
		{\semitransp[25]{
		\item Software
		\begin{itemize}{\semitransp[25]{
			\item Modules
			\item Acquiring software
			\item Compilers
			}}
		\end{itemize}
		\item Managing user jobs
		\begin{itemize}{\semitransp[25]{
			\item Job scheduler
			\item Using SLURM
			\item Partition layout			
			\item Submitting jobs (Examples)
			}}
		\end{itemize}			
		}}
	\end{itemize}
\end{frame}

@@

\begin{frame}
	\frametitle{Globus}\footnotesize
If you already use Globus, the endpoints for the Cray CS300 are:
\begin{itemize}
\item hawaii\#UHHPC1
\item hawaii\#UHHPC1
\end{itemize}
~\\ 
For those that have not used Globus, and wish to try it out, you can learn how to use Globus by visiting the Cyberinfrastructure website for more information:~\\
{\footnotesize \href{http://www.hawaii.edu/its/ci/hpc-resources/hpc-tutorials/globus-quick-start-guide/}{http://www.hawaii.edu/its/ci/hpc-resources/hpc-tutorials/globus-quick-start-guide/}}
\end{frame}

\begin{frame}
	\frametitle{Overview -- Cluster Interaction}
	\begin{itemize}
		\item {\semitransp[25]{Connecting to a cluster @ UH}}
			\begin{itemize}{\semitransp[25]{
			\item Login to the cluster
			\item Verify user permissions
			}}
			\end{itemize}
   {\semitransp[25]{ \item User directories
		\item Transferring files
		\begin{itemize}{\semitransp[25]{
			\item Globus
		}}\end{itemize}
		}}
		\item Software
		\begin{itemize}
			\item Modules
			\item Acquiring software
			\item Compilers
		\end{itemize}
		{\semitransp[25]{
		\item Managing user jobs
		\begin{itemize}{\semitransp[25]{
			\item Job scheduler
			\item Using SLURM
			\item Partition layout
			\item Submitting jobs (Examples)
			}}
		\end{itemize}			
		}}
	\end{itemize}
\end{frame}

@@

\begin{frame}
	\frametitle{Overview -- Cluster Interaction}
	\begin{itemize}
		\item {\semitransp[25]{Connecting to a cluster @ UH}}
			\begin{itemize}{\semitransp[25]{
			\item Login to the cluster
			\item Verify user permissions
			}}
			\end{itemize}
   {\semitransp[25]{ \item User directories
		\item Transferring files
		\begin{itemize}{\semitransp[25]{
			\item Globus
		}}\end{itemize}
		\item Software
		\begin{itemize}{\semitransp[25]{
			\item Modules
			\item Acquiring software
			\item Compilers
			}}
		\end{itemize}
		}}
		\item Managing user jobs
		\begin{itemize}
			\item Job scheduler
			\item Using SLURM
			\item Partition layout
			\item Submitting jobs (Examples)
		\end{itemize}		
	\end{itemize}
\end{frame}

@@

\subsection{Using SLURM}
\begin{frame}
\frametitle{How to use SLURM}
SLURM has a series of commands, each of which allow users to interact with the job scheduler
\begin{itemize}
	\item sbatch -- 
	\item srun --
	\item srun.x11 --
	\item sacct --
	\item scontrol --
	\item scancel --
\end{itemize}  
~\\
Let us look at how the nodes are partitioned, as well as some examples of submitting jobs and create job scripts
\end{frame}

\subsection{SLURM Partitions}
\begin{frame}
\frametitle{Partition layout}
\end{frame}

@@

\section{Frequently Asked Questions}
\begin{frame}
\frametitle{FAQ}
\begin{itemize}
	\item Are files on the cluster backed up?
	\begin{itemize}\footnotesize
	\item \textbf{Answer:} \textbf{NO!}  User files on the cluster \textbf{\emph{are not backed up}}.  It is up to you as the user to validate and maintain your own backups.  The Cyberinfrastructure team takes no responsibility for any data that is lost due to human or mechanical error.  
	\end{itemize}
\end{itemize}
\end{frame}

>>>>

\newcommand{\ci}{Cyberinfrastructure}

@@

\author{Sean Cleveland Ph.D, Ron Merrill Ph.D,\\David Schanzenbach M.S.} % Your name
\institute[University of {\hawaii} -- ITS--CI] % Your institution as it will appear on the bottom of every slide, may be shorthand to save space
{
Information Technology Services \\
Department of {\ci} \\
University of {\hawaii} \\ % Your institution for the title page
\medskip
\textbf{\url{https://www.hawaii.edu/its/ci/}}\\
\textbf{\textit{uh-hpc-help@lists.hawaii.edu}} % Your email address
}
\date{\today} % Date, can be changed to a custom date

@@

\begin{frame}
	\frametitle{Overview -- Cluster Interaction}
	\begin{itemize}
		\item{Connecting to a cluster @ UH}
			\begin{itemize}
			\item Login to the cluster
			\item Verify user permissions
			\end{itemize}
    {\semitransp[25]{\item User directories
		\item Transferring files
		\begin{itemize}{\semitransp[25]{
			\item Globus
		}}\end{itemize}
		\item Software
		\begin{itemize}{\semitransp[25]{
			\item Modules
			\item Acquiring software
			\item Compiliers
			}}
		\end{itemize}
		\item Managing user jobs
		\begin{itemize}{\semitransp[25]{
			\item Job scheduler
			\item Using SLURM
			\item Partitions			
			\item Submitting jobs (Examples)
			}}
		\end{itemize}			
		}}
	\end{itemize}
\end{frame}

@@

\begin{frame}
	\frametitle{Overview -- Cluster Interaction}
	\begin{itemize}
		\item {\semitransp[25]{Connecting to a cluster @ UH}}
			\begin{itemize}{\semitransp[25]{
			\item Login to the cluster
			\item Verify user permissions
			}}
			\end{itemize}
    \item User directories
		{\semitransp[25]{\item Transferring files
		\begin{itemize}{\semitransp[25]{
			\item Globus
		}}\end{itemize}
		\item Software
		\begin{itemize}{\semitransp[25]{
			\item Modules
			\item Acquiring software
			\item Compilers
			}}
		\end{itemize}
		\item Managing user jobs
		\begin{itemize}{\semitransp[25]{
			\item Job scheduler
			\item Using SLURM
			\item Partitions			
			\item Submitting jobs (Examples)
			}}
		\end{itemize}			
		}}
	\end{itemize}
\end{frame}

@@

\begin{frame}
	\frametitle{Overview -- Cluster Interaction}
	\begin{itemize}
		\item {\semitransp[25]{Connecting to a cluster @ UH}}
			\begin{itemize}{\semitransp[25]{
			\item Login to the cluster
			\item Verify user permissions
			}}
			\end{itemize}
    {\semitransp[25]{\item User directories}}
		\item Transferring files
		\begin{itemize}
			\item Globus
		\end{itemize}
		{\semitransp[25]{
		\item Software
		\begin{itemize}{\semitransp[25]{
			\item Modules
			\item Acquiring software
			\item Compilers
			}}
		\end{itemize}
		\item Managing user jobs
		\begin{itemize}{\semitransp[25]{
			\item Job scheduler
			\item Using SLURM
			\item Partitions			
			\item Submitting jobs (Examples)
			}}
		\end{itemize}			
		}}
	\end{itemize}
\end{frame}

@@

\begin{frame}
	\frametitle{Globus}\footnotesize
If you already use Globus, the endpoints for the Cray CS300 are:
\begin{itemize}
\item hawaii\#UHHPC1
\item hawaii\#UHHPC1
\end{itemize}
~\\ 
For those that have not used Globus, and wish to try it out, you can learn how to use Globus by visiting the {\ci} website for more information:~\\
{\footnotesize \href{http://www.hawaii.edu/its/ci/hpc-resources/hpc-tutorials/globus-quick-start-guide/}{http://www.hawaii.edu/its/ci/hpc-resources/hpc-tutorials/globus-quick-start-guide/}}
\end{frame}

\begin{frame}
	\frametitle{Overview -- Cluster Interaction}
	\begin{itemize}
		\item {\semitransp[25]{Connecting to a cluster @ UH}}
			\begin{itemize}{\semitransp[25]{
			\item Login to the cluster
			\item Verify user permissions
			}}
			\end{itemize}
   {\semitransp[25]{ \item User directories
		\item Transferring files
		\begin{itemize}{\semitransp[25]{
			\item Globus
		}}\end{itemize}
		}}
		\item Software
		\begin{itemize}
			\item Modules
			\item Acquiring software
			\item Compilers
		\end{itemize}
		{\semitransp[25]{
		\item Managing user jobs
		\begin{itemize}{\semitransp[25]{
			\item Job scheduler
			\item Using SLURM
			\item Partitions
			\item Submitting jobs (Examples)
			}}
		\end{itemize}			
		}}
	\end{itemize}
\end{frame}

@@

\begin{frame}
	\frametitle{Overview -- Cluster Interaction}
	\begin{itemize}
		\item {\semitransp[25]{Connecting to a cluster @ UH}}
			\begin{itemize}{\semitransp[25]{
			\item Login to the cluster
			\item Verify user permissions
			}}
			\end{itemize}
   {\semitransp[25]{ \item User directories
		\item Transferring files
		\begin{itemize}{\semitransp[25]{
			\item Globus
		}}\end{itemize}
		\item Software
		\begin{itemize}{\semitransp[25]{
			\item Modules
			\item Acquiring software
			\item Compilers
			}}
		\end{itemize}
		}}
		\item Managing user jobs
		\begin{itemize}
			\item Job scheduler
			\item Using SLURM
			\item Partitions
			\item Submitting jobs (Examples)
		\end{itemize}		
	\end{itemize}
\end{frame}

@@

\subsection{SLURM}
\begin{frame}
\frametitle{SLURM}
\begin{block}{How are jobs scheduled?}
User submitted jobs are assigned a priority using a fairshare algorithm.  Factors such as: runtime, resource usage request, age of job, and amount of compute time a user has used in recent history are all factored in to assigning a priority to a give job.   
\end{block}
\end{frame}

\begin{frame}
\frametitle{How to use SLURM}
SLURM has a series of commands, each of which allow users to interact with the job scheduler
\begin{itemize}
	\item sbatch -- 
	\item srun --
	\item srun.x11 --
	\item sacct --
	\item scontrol --
	\item scancel --
\end{itemize}  
~\\
Let us look at how the nodes are partitioned, as well as some examples of submitting jobs and create job scripts
\end{frame}

\subsection{SLURM Partitions}
\begin{frame}
\footnotesize
\frametitle{Partitions}
\begin{block}{What is a partition?}\footnotesize
A partition can be thought of as a group of nodes/resources divided into possibly overlapping sets.  Each partition can be considered as a job queue, each of which has an assortment of constraints such as job size limit, job time limit, users permitted to use it, etc. Priority-ordered jobs are allocated nodes within a partition until the resources (nodes, processors, memory, etc.) within that partition are exhausted.\footnotemark
\end{block}
\begin{itemize}
	\item The Cray CS300 currently has six public partitions:~\\\textbf{lm.q}, \textbf{community.q}, \textbf{exclusive.q}, \textbf{sb.q}, \textbf{kill.q}, \textbf{htc.q}
	\item Jobs submitted to kill.q and htc.q can be preempted by jobs in other partitions
	\item More details on partition constraints can be viewed by visiting the \href{http://www.hawaii.edu/its/ci/hpc-resources/slurm-partitions/}{{\ci} website} or by using the following command
	\begin{itemize}\footnotesize
		\item[] \texttt{$[$login \ctilde$]$\$ scontrol show partition $<$partition name$>$}
		\item[] Example: \texttt{$[$login \ctilde$]$\$ scontrol show partition community.q}
	\end{itemize}

\end{itemize}
\footnotetext[2]{\tiny \url{http://slurm.schedmd.com/quickstart.html}}
\end{frame}

\begin{frame}
\frametitle{Partitions}
\includegraphics[width=0.95\textwidth]{images/partitions}
\end{frame}

@@

\section{Frequently Asked Questions}
\begin{frame}
\frametitle{FAQ}
\begin{itemize}
	\item Are files on the cluster backed up?
	\begin{itemize}\footnotesize
	\item \textbf{Answer:} \textbf{NO!}  User files on the cluster \textbf{\emph{are not backed up}}.  It is up to you as the user to validate and maintain your own backups.  The {\ci} team takes no responsibility for any data that is lost due to human or mechanical error.  
	\end{itemize}
\end{itemize}
\end{frame}
